---
title: "Almonds_of_Wrath"
author: "Koshlan"
date: "August 5, 2015"
output: html_document
---
### Spatial Vignette 1: The Nuts of Wrath: Kern County, CA 

As an environmental scientist, 
I am fascinated by spatial data, natural resource use, and
the consequences of human activities on environmental quality. 

In this vignette, I explore the expansion of almonds
in Kern Country, California from 2010 to 2014. 

The expansion of almonds and other high-value woody perennial crops (e.g.  
such as pistachios and walnuts) has been well documented by local and 
national media covering the California drought; however, I am interested in some basic land-use questions:

* First, what is the plausible **net effect** on irrigation of nut tree expansion.
Obviouslly these crops need water, but what were the prior-uses and 
associated water demand of the lands converted to new Almond orchards. 

* Second, what associations may exist between agricultural conversions 
and pesticide use. 

* Finally what portions of Kern county are most likely to be further converted to 
future nut orchards.

```{r, packages, message=F, warning=F}
library(raster)
library(maptools)
```

### Where to begin?

Kern County publishes permited agricultural boundaries. This data is provided as a shapefile. We can read the polygon file in using maptoos::readShapePoly; however the object lacks information about the coordinate systems and projection, two essential
aspects for contextualizing spatial data. For a quick reference on this, check out

*https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf

```{r, boundaries }
kern_boundaries <- maptools::readShapePoly("/Users/koshlan/Dropbox/KernCounty/kern2014/kern2014.shp")
```

This information is provided on at the Kern County Department Of Agriculture And Measurement Standards (http://www.kernag.com/gis/gis-data.asp). 

* **NAD_1983_StatePlane_California_V_FIPS_0405_Feet**
* Projection:  Lambert_Conformal_Conic
* False_Easting:  6561666.66666667
* False_Northing:  1640416.66666667
* Central_Meridian:  -118.00000000
* Standard_Parallel_1:	34.03333333
* Standard_Parallel_2:	35.46666667
* Latitude_Of_Origin:	33.50000000
* Linear Unit:	Foot_US
* Geographic Coordinate System:	GCS_North_American_1983
* Datum:	D_North_American_1983
* Prime Meridian:	Greenwich
* Angular Unit:	Degree

The projection
is **NAD_1983_StatePlane_California_V_FIPS_0405_Feet**, and to work with it in R
we need the abbreviated string (proj4string). I took my guess and looked it 
up at:

* http://epsg.io/102645. 

We can then assign this projection to the polygons object
```{r, assign_projection}
sp::proj4string(kern_boundaries) <- CRS("+proj=lcc +lat_1=34.03333333333333 +lat_2=35.46666666666667 +lat_0=33.5 +lon_0=-118 +x_0=2000000 +y_0=500000.0000000002 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192")
```

Here I check that our boundary files lands squarely in California.

```{r} 
states <- readShapePoly("/Users/koshlan/Dropbox/KernCounty/cb_2013_us_state_20m/cb_2013_us_state_20m.shp", proj4string=CRS("+proj=longlat"))
california <- states[states$STATEFP == "06",]
plot(california)
check_kern = sp::spTransform(kern_boundaries, proj4string(california))
```

### Crop Data

The national agricultural statistics service generates a 30 meter resolution estimate of crop-cover, which is based on a decision tree classifier from landsat imagery. This data is in .geotiff format. If our projection is correctly assigned
the Kern County crop boundaries should correspond with cropland in the raster dataset. Once we have assigned a projection to a polygon object we can tranform it to another projection. 

```{r}
# geotiff crop data from 2 years in Kern County, CA  (USDA Cropscape)
kern2010 <- raster("/Users/koshlan/Dropbox/KernCounty/CDL_2010_06029.tif")
kern2014 <- raster("/Users/koshlan/Dropbox/KernCounty/CDL_2014_06029.tif")
kern_boundaries.t <- spTransform(kern_boundaries, proj4string(kern2010))
plot(kern2014)
plot(kern_boundaries.t, add = T, border = "white", lwd = 0.1)
```
