---
title: "Analysis_of_Polygons"
author: "Koshlan"
date: "July 15, 2015"
output: html_document
---


To produce more reproducible research and get the most out of reading, I am 
migrating tutorial notes to Rmarkdown using knitr.

This particular document was created after following a tutorial from Brunsdon and Comber's 
(2015) An Introduction to R for Spatial Analysis. 

Foremost, they had very nice vizualization of spatial adjacency networks.

### Packages
Here are the preliminaries:
```{r, message=F, warning=F}
require(GISTools)
require(SpatialEpi)
require(rgdal)
require(spdep)
```

The dataset is loaded, coordinates transformed, and a descriptive variable extracted.
From which we can plot spatial distribution of smoking in Pennsylvania:
```{r}
data(pennLC)
#str(pennLC)
penn.state.latlong <- pennLC$spatial.polygon
penn.state.utm <- spTransform(penn.state.latlong, 
                              CRS("+init=epsg:3724 +units=km +ellps=WGS84"))
smk<-pennLC$smoking$smoking *100
```

```{r, echo=FALSE, out.width='350px', dpi=200}
shades = auto.shading(smk, 
                      n =6, 
                      cols = brewer.pal(5, "Reds"))

par(mar =c(0,0,0,0))
choropleth(penn.state.utm, smk, shades)
choro.legend(538.336,
             4394,
             shades, 
             title = '% Smoking',
             cex = .5)

```
### Polygons Network
But, what I really liked in the tutorial was relationship vizualization:
```{r, echo=TRUE, out.width='350px', dpi=200}
penn.state.nb <- poly2nb(penn.state.utm)
# The object class "nb" stores a list of neighboring polygons
plot(penn.state.utm, border = "lightgrey")
plot(penn.state.nb, coordinates(penn.state.utm), add = T, col = "#0000FF50", lwd = 2)
```

### Lagged Means
The authors suggest a first order exploratory approach:
analysis of lagged means, appropriate to continuous data.

That is, the value of polygon divided vs. the unweighted mean of its neighbors. 
A comparison of the polygon value to that of its lagged mean shows anomolies 
that differ from surroundings, as well as points on the 45 degree line that 
share a value similar  to their local neighborhood. 

Moreover, an enrichment of points in 1st and 3rd quadrants 
manifests a degree of spatial auto correlation.

```{r, echo=TRUE}
# This object holds the (i) neighbors of each polygon and (ii) their weights

penn.state.lw <- nb2listw(penn.state.nb)
smk.lagged.means <- lag.listw(penn.state.lw, smk)
par(mar =c(5,5,5,5))
plot(smk, smk.lagged.means, asp=1, xlim = range(smk), ylim = range(smk))
abline(0,1)
abline(v=mean(smk), lty=2)
abline(h =mean(smk.lagged.means), lty =2)
```

### Moran's I
Moran's I is a measure of spatial autocorrelation amoung polygons in a spatial domain.

The starting point is a matrix (W) that relates the degree of dependence between
the ith and jth polygon, and the statistic follows:

 $\begin{equation}
I = \frac{n}{\displaystyle\sum_{i=1}^n \sum_{j=1}^n w_{ij}}
\frac{\displaystyle\sum_{i=1}^n \sum_{j=1}^n w_{ij}(x_i - \bar{x})(x_j -
  \bar{x})}{\displaystyle\sum_{i=1}^n (x_i - \bar{x})^2},\label{eq:morani}
\end{equation}$

The moran test can be run from spdep no the listw object
```{r, echo=TRUE}
require(spdep)
moran.test(smk,penn.state.lw)
```

Brunsdon and Comber's make two noteworthy observations. First, 
When the W matrix is standardized so the rows equal one, when the weights are 
exactly those used in a lagged mean test. Second, the Moran's I coefficient is 
dimensionless number that varies with valuses of W matrix. Thus, it's interpretation should involve
a comparison hus should be compared to a null hypothesis that there was 
no spatial correlation at all and the value could be achieved by random chance.

Three approaches are offered.

1. Parametric assume the value are drawn from a gaussian, with a test statistic (see page 233)
2. Premuation test, with convergens to a normal. That is, premute all the values randomly amoung polygons.
3. Monte Carlo

Note the two choices

```{r, echo=TRUE}
require(spdep)
moran.test(smk,penn.state.lw, randomisation = TRUE)
moran.test(smk,penn.state.lw, randomisation = FALSE)
moran.mc(smk,penn.state.lw, 1000)
```


### Spatial Autocorrelation

####SAR MODEL

Weights are often based on polygons adjacency. Recall that in Kriging, a similar variance 
covariance matrix is based on physical distance.




### Summary. 

So what are the take aways. To work with polygon spdep is the package with the glue.
A crucial set of objects for encapsilating information about polygon adjacency are nb and listw.

A function for determining nieghbors

spdep::poly2nb()

A function for creating an object to hold neighbors and weights 

spdep::nb2listw()



Some things left to discover:

What about weights based on centroid distance or edge distance?

What about neighbors of neighbors?

How scalable are these objects and functions?

Identifying spatial correlation is easy? How can it be used to improve prediction.


















